Continuous estimates of heat emission at Mt. Ruapehu using the Unscented Kalman Smoother

### 2021-07-16 manuscript/manuscript.tex
document compiles with all sections added

Increases in heat and steam input may be caused by the ascent of magma or
changes in the permeability of the hydrothermal system and are, therefore,
important parameters to be monitored. First attempts to quantify the amount of
heat and steam entering the lake date back to \citep{Dibble1966}. Over the past
decade water temperature and lake level have been monitored continuously
and water samples for ion concentration measurements have been taken roughly
every two to three months. This enables us to continuously estimate heat and
steam input into the lake and, in principle, automatically detect changes in
these variables as soon as they occur.

Inference of the heat and steam input is typically done through a mass and
energy balance calculation: the sum of mass and energy flow into and out of the
lake has to account for the observed changes in lake temperature, water level
and ion dilution \citep[e.g.,][]{Hurst1981, Hurst1991, Stevenson1992,
Fournier2009, Scott1994}. At Crater Lake, and probably at most other volcanic
lakes, this is an underdetermined problem, that is the number of unknowns
exceeds the number of constraints. Hence, it is necessary to constrain some of
the parameters in the mass and energy balance model using prior knowledge. Further,
some observations and input parameters come with large uncertainties and are sampled
irregularly.

\section*{Section title}
Text for this section \ldots
\subsection*{Sub-heading for section}
Text for this sub-heading \ldots
\subsubsection*{Sub-sub heading for section}
Text for this sub-sub-heading \ldots
\paragraph*{Sub-sub-sub heading for section}
Text for this sub-sub-sub-heading \ldots
In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
addition, we examine a common modeling assumption and note the
importance of considering the tails of the extinction time $T_x$ in
studies of escape dynamics.
We will first consider the expected resistant population at $vT_x$ for
some $v>0$, (and temporarily assume $\alpha=0$)
%
\[
 E \bigl[Z_1(vT_x) \bigr]= E
\biggl[\mu T_x\int_0^{v\wedge
1}Z_0(uT_x)
\exp \bigl(\lambda_1T_x(v-u) \bigr)\,du \biggr].
\]
%
If we assume that sensitive cells follow a deterministic decay
$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
estimate the expected value as
%
\begin{eqnarray}\label{eqexpmuts}
E\bigl[Z_1(vT_x)\bigr] &=& \frac{\mu}{r}\log x
\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du
\nonumber\\
&=& \frac{\mu}{r}x^{1-{\lambda_1}/{\lambda_0}v}\log x\int_0^{v\wedge
1}x^{-u(1+{\lambda_1}/{r})}\,du
\nonumber\\
&=& \frac{\mu}{\lambda_1-\lambda_0}x^{1+{\lambda_1}/{r}v} \biggl(1-\exp \biggl[-(v\wedge1) \biggl(1+
\frac{\lambda_1}{r}\biggr)\log x \biggr] \biggr).
\end{eqnarray}
%
Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,khar,zvai,xjon,marg}).
%\nocite{oreg,schn,pond,smith,marg,hunn,advi,koha,mouse}

>>>>

Increases in heat input may be caused by the ascent of magma or
changes in the permeability of the hydrothermal system and are, therefore,
important parameters to be monitored. First attempts to quantify the amount of
heat entering the lake date back to \citep{Dibble1966}. Since 2016
water temperature and lake level have been monitored continuously
and water samples for ion concentration measurements have been taken roughly
every two to three months. This enables us to continuously estimate heat input
into the lake and, in principle, automatically detect changes as soon as they occur.

Inference of the heat input is typically done through a mass and
energy balance calculation: the sum of mass and energy flow into and out of the
lake has to account for the observed changes in lake temperature, water level
and ion dilution \citep[e.g.,][]{Hurst1981, Hurst1991, Stevenson1992,
Fournier2009, Scott1994}. At Crater Lake, and probably at most other volcanic
lakes, this is an underdetermined problem, that is the number of unknowns
exceeds the number of constraints. Hence, it is necessary to constrain some of
the parameters in the mass and energy balance model using prior knowledge. Further,
some observations and input parameters come with large uncertainties and are sampled
irregularly.

In the following, we propose a probabilistic (Bayesian) inference approach which
naturally handles these difficulties and transparently propagates uncertainties of
observations and prior assumptions into posterior uncertainties of heat input.

\begin{figure}
\includegraphics[width=\textwidth]{figures/overview.png}  
\caption{Image of Mt. Ruapehu's crater lake from 2008 (\textcopyright Karen
	Britten, GNS Science). The red O marks the location of the lake outlet. The
	location of Mt. Ruapehu is shown as a red triangle on the insert map.}
  \label{overview}
\end{figure}

\section*{Probabilistic inference method}\label{Pim}

\subsection*{Physical model}\label{phm}

Following \cite{Hurst1991} and \cite{Stevenson1992} the mass and energy balance
can be written in terms of three coupled ordinary differential equations. The
rate of lake water mass ($M$) change can be described as:
\begin{equation}\label{ode_M}
	\frac{dM}{dt} = \dot{M_s} - \dot{M_o} - \dot{M_e} + \dot{M_i}
\end{equation}

Where $\dot{M_s}$ is the rate of water inflow through melt and precipitation,
$\dot{M_o}$
is the outflow rate through seepage and overflow, $\dot{M_e}$ is the rate of mass
loss through evaporation and $\dot{M_i}$ is the rate of steam input through the vent
system beneath the lake.

The energy content of the lake can be described with the notion of
enthalpy, $H$. At constant pressure, enthalpy is defined as:
\begin{equation}
	H(T) = H(T_0) + \int_{T_0}^{T}c_pM\tau d\tau
\end{equation}

where $T$ is temperature and $c_p$ is the specific heat. For all following
equations we take $H(T_0=0)$ as our reference enthalpy and $c_p=c_w=constant$,
with $c_w$ the specific heat of water, such that the enthalpy of lake water at
temperature $T$ can be written as $H(T) = c_wMT$. Because Ruapehu Crater Lake is
 at an altitude of $\sim$~2500~m and big parts of its catchment are covered by snow
and ice all year round we assume the temperature of incoming water to be at 0
$^{\circ}C$.

The rate of enthalpy change of the lake water can thus be written as:
\begin{equation}\label{ode_E}
    \frac{dH}{dt}=\frac{d(c_wTM)}{dt}=c_wT\dot{M} + c_wM\dot{T} =
	-\dot{Q_e} - c_wT\dot{M_o} + \dot{Q_{eff}} + \dot{Q_r}
\end{equation}

where $T$ is the lake water temperature, $\dot{Q_r}$ is
the rate of energy gain due to solar irradiation (short-wavelength radiation),
and $c_wT\dot{M_o}$ is the rate of energy loss through outflow/seepage. $\dot{Q_e}$ is
the rate of surface energy losses due to evaporation (forced and free
convection), sensible heat, and long-wavelength radiation (net loss). Several
empirical and theoretical equations exist to describe the surface losses and
both \cite{Stevenson1992} and \cite{hurstCraterLakeEnergy2015} discuss them in detail. We
follow here mostly the choice of \cite{hurstCraterLakeEnergy2015} as it includes the recent
development by \cite{sartoriCriticalReviewEquations2000} on forced (i.e. wind-induced) convection
(see Appendix \ref{A} for more details). $\dot{Q_{eff}}$ is the effective rate of
energy gain from volcanic sources after heating incoming water:
\begin{equation}
	\dot{Q_{eff}}=\dot{Q_i} - c_wT\dot{M_s}
\end{equation}
with $\dot{Q_i}$ the rate of energy input from volcanic sources.

Following from Equation \ref{ode_E} the temperature change $\frac{dT}{dt}$ can then be written as:
\begin{equation}\label{ode_T}
	\frac{dT}{dt}=\frac{1}{c_wM}\left(-\dot{Q_e} - c_wT\dot{M_o} + \dot{Q_{eff}} + \dot{Q_r}\right )
	-\frac{T}{M}\dot{M}
\end{equation}

Finally, the change in ion concentration of, for example, $Mg^{2+}$ can be expressed as:
\begin{equation}\label{ode_X}
	\frac{dX}{dt}=-\dot{M_o}\frac{X}{M}
\end{equation}

Where $X$ is the total ion amount in the lake. This equation is only true under
the assumption that $X$ only changes due to dilution and there is no re-supply
between eruptive episodes. Equation \ref{ode_X} gives an estimate of the average
outflow over periods on the order of months but does not capture shorter term
variations in the outflow.

In a sensitivity analysis for $\dot{T}$ (Figure \ref{sens}) we can see a linear
relationship with respect to $\dot{Q_i}$ and $\dot{M_o}$ and a mildly
non-linear relationship with respect to $\dot{Q_i}$ and current season
(note that this assumes southern hemisphere seasons). The relationship with respect
to $T$ and windspeed, however, is strongly non-linear. This is because $\dot{Q_e}$
and $\dot{M_e}$ are non-linear functions of windspeed and $\dot{Q_i}$. This further 
complicates the inference of $\dot{Q_i}$.

\begin{figure}
    \includegraphics[width=\textwidth]{figures/sensitivity_analysis.png}  
    \caption{Sensitivity analysis of $\dot{T}$ with respect to $\dot{Q_i}$ 
             and $\dot{M_o}$ (top panel), $\dot{Q_i}$ and season 
             (center panel), and $T$ and windspeed (bottom panel). The model
              parameters that have been held fixed are described in each 
              panel's title. Warmer colors indicate a temperature increase
              ($\dot{T} > 0$) and cooler colors a temperature decrease
              ($\dot{T} < 0$)}
      \label{sens}
\end{figure}

\subsection*{Inference}\label{inf}

The objective of the inference problem is to estimate the rate at which heat
is entering the lake from the hydrothermal system below.
To constrain this we use the currently available observations of $T$, $M$,
$X$, and $\dot{M_o}$ as well as numerical weather model data on windspeed.
Putting it more concisely, we are looking for

\begin{equation}\label{bsm}
    p(\bm{x}_k|\bm{y}_{1:t})
\end{equation}

where $\bm{y}_{1:t}$ is the time series of $m$ observations and $x_k$ comprises
the $n$ model parameters described in Section \ref{phm} at time step $k$. Note
that $x_k$ also includes $T$, $M$, $\dot{M_o}$, and $X$. We distinguish here
between the state of the lake and observations of this state
($T$, $M$, $X$, $\dot{M_o}$) which are to some degree uncertain.

The following equations are based on \cite{sarkkaBayesianFilteringSmoothing2013}
which we refer the reader to for more proofs and detailed explanations.

Equation \ref{bsm} is the marginal probability of $\bm{x}_k$ given all
observations until time $t$. We assume that the states of the lake form a
Markov chain, meaning that the state of the lake at time step $k$, $\bm{x}_k$,
only depends on the state at adjacent time steps:

\begin{equation}\label{markov1}
    p(\bm{x}_k|\bm{x}_{1:k-1}, \bm{y}_{1:k-1}) = p(\bm{x}_k|\bm{x}_{k-1})
\end{equation}

and

\begin{equation}\label{markov2}
    p(\bm{x}_{k-1}|\bm{x}_{k:t}, \bm{y}_{k:t}) = p(\bm{x}_{k-1}|\bm{x}_{k})
\end{equation}

With this assumption and using \textit{Bayes' rule}, Equation \ref{bsm} can be
rewritten as:

\begin{equation}\label{bsm1}
p(\bm{x}_k|\bm{y}_{1:T}) = p(\bm{x}_k|\bm{y}_{1:k})
    \int\left[\frac{p(\bm{x}_{k+1}|\bm{x}_k)p(\bm{x}_{k+1}|\bm{y}_{1:T})}{p(\bm{x}_{k+1}|\bm{y}_{1:k})}\right]d\bm{x}_{k+1}
\end{equation}

with

\begin{eqnarray}\label{bfilt1}
p(\bm{x}_{k+1}|\bm{y}_{1:k}) & = & \int p(\bm{x}_{k+1}|\bm{x}_k)p(\bm{x}_k|\bm{y}_{1:k})d\bm{x}_k \\
        \label{bfilt}
    p(\bm{x}_k|\bm{y}_{1:k}) & = & \frac{1}{z_k} p(\bm{y}_k|\bm{x}_k)p(\bm{x}_k|\bm{y}_{1:k-1}) \\
        \label{zetk}
    z_k & = & \int p(\bm{y}_k|\bm{x}_k) p(\bm{x}_k| \bm{y}_{1:k-1}) d\bm{x}_k  
\end{eqnarray}

using the \textit{Chapman-Kolmogorov} equation we can write:

\begin{equation}
p(\bm{x}_k|\bm{y}_{1:k-1}) = \int p(\bm{x}_k|\bm{x}_{k-1})p(\bm{x}_{k-1}|\bm{y}_{1:k-1})d\bm{x}_{k-1} 
\end{equation}

Equations \ref{bsm1} to \ref{zetk} describe a recursive way to compute Equation
\ref{bsm} in which we only need to know $\bm{x}_1$,
$p(\bm{x}_k|\bm{x}_{k-1})$, and $p(\bm{y}_k|\bm{x}_k)$. Equation \ref{bsm1} and
\ref{bfilt} are also known as the Bayesian Smoothing and Bayesian Filtering
equation, respectively. If $p(\bm{x}_k|\bm{x}_{k-1})$ and
$p(\bm{y}_k|\bm{x}_k)$ are linear transformations with normally distributed
errors, the \textit{Kalman filter} \cite{Kalman1960} and Kalman smoother
\cite{Rauch1965} represent closed form solutions. In our case, these
transformations are:

\begin{eqnarray}
p(\bm{x}_k|\bm{x}_{k-1}) & = & f(\bm{x}_{k-1}) + \bm{q}_{k-1} \\
    p(\bm{y}_k|\bm{x}_k) & = & \bm{x}_k[1:m] + \bm{r}_k   
\end{eqnarray}

where $f(\bm{x}_{k-1})$ is the physical model described in Section \ref{phm}
which is strongly non-linear as shown in Figure \ref{sens}. If we assume the errors, $\bm{r}_k$ and
$\bm{q}_{k-1}$, to be normally distributed ($\bm{q}_{k-1} \sim N(\bm{0},
\bm{Q}_{k-1})$, $\bm{r}_k \sim N(\bm{0}, \bm{R}_k)$), we can compute an
approximate solution to \ref{bsm1} using the \textit{Unscented Kalman Filter}
\citep{Merwe2004}.

We estimate $\bm{R}_k$ from the daily variations in the observations but we do
not know $\bm{x}_1$ and $\bm{Q}_{k-1}$. Setting $\bm{x}_1$ to an arbitrary value
within physically reasonable bounds but large uncertainty works well in practice.
While $\bm{Q}_{k}$ can be formally optimised we treat it here as a design
parameter that we modify such that the observations are matched within their
error bounds and the inferred heat input rate varies smoothly. The latter is
motivated by the fact that the physical model described in \ref{phm} is not
able to model short-term transient changes in lake volume and temperature.
These can be caused by rain or increased meltwater influx and the assumption
of a perfectly mixed lake typically does not apply in these situations.

\section*{Results}
\subsection*{Synthetic test}\label{syn_test}
To evaluate the inference scheme we design a synthetic test, consisting of a very
large cylinder, the same volume as Ruapehu Crater Lake. We keep the windspeed
and surface water inflow constant at 4.5 m/s and 10 kt/day, respectively. To
model the outflow we assume the outlet to be a pipe 0.2 $m^2$ in cross section
and then use Bernoulli's equation to model the water outflow based on the water
level in the cylinder above the pipe. We assume that the incoming heat follows
a combination of a third-order polynomial and a sinusoid function and that the
enthalpy of the incoming steam is 2.8 which corresponds to the enthalpy of
saturated steam at hydrostatic pressures similar to those at the bottom of the
lake \cite{Mayhew1978}. To the resulting synthetic observations we add
uncertainties similar to those estimated from real observations (see Section
\ref{realdata}).

Figure \ref{syn_example} shows the result of the probabilistic inference of
heat input compared to the true input of the synthetic model. The observations
are well fit within their error bounds but we see some descrepancies between
the original and the recovered heat input rates, especially for very steep
changes in the input rate. This is the result of preferring a smooth solution
as already mentioned in Section \ref{inf}.

It is noteworthy that the mass influx and outflux rates are recovered
quite accurately despite having no data in the former and very little data
in the latter case.

\begin{figure}
	\includegraphics[width=\textwidth]{figures/synthetic_inversion_uks.png}  
    \caption{Inversion result for the synthetic test described in Section 
        \ref{syn_test}. Turquoise lines and markers represent hypothetical
        observations; red dashed lines are the inversion results and black
        dashed lines are the true input to the synthetic test.}
  \label{syn_example}
\end{figure}

\subsection*{Real data}\label{realdata}
Temperature at RCL is measured by a integrated-circuit temperature device
(Texas Instruments LM35) located roughly 1.9 m below the lake's current
overflow level. The water level sensor is a hydrostatic pressure sensor (First
Sensor KTE/KTU/KTW6000\ldots CS Series) and it is co-located with the
temperature sensor. Both sensors have a sampling interval of 15 minutes and
transmit data several times a day via satellite to GNS Science, which monitors
New Zealand's geohazards through its GeoNet project (www.geonet.org.nz
 last accessed 16/07/2021).
Water samples are taken 8-10 times a year from which subsequently, among many
other analytes, concentrations of $Mg^{++}$ and $Cl^-$ are determined at GNS
Science's Geothermal Analytical Laboratory. We will focus here on $Mg^{++}$
concentration as it shows less deviation from the assumption of constant
dilution than $Cl^-$.

To estimate uncertainties of temperature and water level observations, we 
compute their daily averages and standard deviations. Uncertainties of $Mg^{++}$
concentrations are estimated directly when two samples were taken on the same day.
Otherwise, their uncertainty is taken to be the median of directly estimated
uncertainties. $\dot{M_o}$ is measured once or twice per year and we further know
from field observations that the lake is not overflowing beneath a certain water
level.

As there is no weather station close to RCL we assume the wind speed from the
New Zealand Metservice's numerical weather prediction model with a standard
deviation of ??.

\begin{figure}
% 	\includegraphics[width=.9\textwidth]{clemb_inference_result}  
% 	\caption{Inference results for the energy input rate (A),
% 		the steam input rate (B), the water inflow (C) and 
% 		outflow rate (D) for temperature, water level and 
% 		$Mg^{++}$ observations at RCL in 2018.
% 		The color shows an interpolation of 1D kernel density 
% 		estimates (kdes) of the posterior samples for each day's 
% 		inference result. The solid black line marks the expected
% 		energy input rate. (E) shows the Bhattacharyya distance 
% 	        between the probability distributions of the observations
% 	        and the probability distributions of the inference results.}
\label{inference_result}
\end{figure}

\section*{Discussion and conclusion}
Although we use a tighter constraint on the
outflow than just the dilution of $Mg^{++}$, inflow and outflow are still
poorly constraint, especially when the lake level is above overflow level. As we already
showed in the synthetic test in Section \ref{syn_test}, the $MG^{++}$ alone has
too large uncertainties to properly constrain outflow rate. The probabilistic
inference results thus help us determine future priorities for field
observations and work is underway to establish a quantitative relationship
between lake level and outflow rate.

\cite{hurstCraterLakeEnergy2015} and \cite{Stevenson1991} both found that they required very
high values of enthalpy for the incoming steam to avoid violating physical
boundary conditions such as negative values for the inflow rate of meteoric
water. Their explanation was the existence of a heat pipe beneath the lake where
gas and fluid phases coexist. As the rising gas enters the lake relatively cool
lake water enters the vent leading to an overall lower addition of mass to the
lake and resulting in an enthalpy value that is double than expected for
saturated steam at hydrostatic pressures likely to be found at the bottom of the
lake. By including uncertainties in our observations we no longer require the
high enthalpy values to satisfy the physical constraints. In fact, since we \textit{a
priori} only allow physically reasonable parameter ranges we find the
combinations of parameter values that reproduce the observations within their
respective uncertainties.

Inferring the heat input rate into a crater lake is a corner piece in estimating the
total energy budget of a volcanic system as, for example, described by \cite{Adams1989a}
for Po\'as volcano in Costa Rica. Similar to a volcano's gas budget, it can either
indicate migration of magma or changes in permeability within the vent system. To
answer whether it is one or the other requires more detailed numerical modelling
of the magmatic and hydrothermal system. Many studies have demonstrated the 
value of non-linear Kalman Filters for data assimilation with very complex numerical 
models \citep{}. The next logical step, therefore, is to include models of magma
degassing and hydrothermal fluid flow to model the heat input rate, volume, and
enthalpy of the steam entering the lake. Because Kalman Filters return a probabilistic
estimate of model parameters, even crude, first-order models could be used
as long as appropriate model uncertainties are assigned.

* probabilistic approach allows for Bayesian comparison between evaporation models

Although an increased heat input rate at Ruapehu
often goes along with an increase in seismic tremor levels (see Figure ??), this is
not always the case and so it is an independent indicator of potential changes in
the volcano.

Figure ?? shows the comparison of daily averages of volcanic tremor amplitudes
(RSAM; reference), heat input rate and water temperature. We can see that heat
input rate correlates well with tremor values with a correlation coefficient of
?. If we only consider heat pulses above ? MW (see dashed line in Figure ?), the
correlation coefficient increases to ?. Although correlation does not imply
causality it would seem plausible that high heat, and therefore gas, fluxes
could generate volcanic tremor. A mechanism of heat flow generating tremor
has been described by \citet{Iwamura2005}.

In conclusion, the combination of the mass and energy model of \citet{Hurst2015}
and \citet{Stevenson1992} combined with the probabilistic inference method we
developed here provides a powerful method to continuously infer the heat input
rate into Ruapehu Crater Lake. It allows us to combine disparate data streams,
quantify our beliefs in the results and thereby augment our monitoring
capabilities.

\appendix
\section{Surface losses}\label{A}
\subsection{long-wavelength radiation}
The net loss through long-wavelength radiation can be written as:
\begin{equation}
    Q_{rad} = A\sigma(\epsilon_w T_s^4 - \epsilon_a T_a^4)
\end{equation}

where

\begin{tabular}{cp{8cm}}
    A & is the surface of the lake (in $m^2$) \\
    $\sigma$ & is the Stefan-Boltzmann coefficient ($5.67e^{-8} \frac{W}{m^2 K^4}$) \\
    $\epsilon_w$ & is the lake emissivity \\
    $T_s$ & is the water surface temperature (here assumed to be 1 $^{\circ}C$ less than the
    measured temperature) \\
    $\epsilon_a$ & is the effective atmospheric emissivity \\
    $T_a$ & is the effective air temperature (here assumed to be 0.9 $^{\circ}C$) \\ 
\end{tabular}

\subsection{Evaporation}
As proposed by \citet{Hurst2015} we use the equation of \citet{Adams1990} but
replace the term for forced convection by that of \citet{Sartori2000}:

\begin{equation}
    Q_{evap} = [(2.2(T_{sv} - T_{av})^{1/3}(e_s - e_a))^2 + 
    (L(0.00407u^{0.8}X_0^{-0.2})(e_s - e_a)/P_a)^2]^{1/2}
\end{equation}

where

\begin{tabular}{cp{8cm}}
    $T_{sv}$, $T_{av}$ & are the virtual surface and air temperatures respectively.
    This correction accounts for the extra buoyancy of water vapour compared to air
    (see Equation \ref{vtemp} for details) \\
    $e_s$, $e_a$ & are the saturation vapour pressure and the ambient air vapour
    pressure respectively \\
    $L$ & is the latent heat of evaporation \\
    $u$ & is the wind speed \\
    $X_0$ & is the fetch, or characteristic length, of the lake\\
    $P_a$ & is the atmospheric pressure \\
\end{tabular}

\begin{equation}\label{vtemp}
T_{xv} = \frac{T_x}{(1-0.378e_x/P_a)} \qquad \mathrm{for}\quad x = s\ \mathrm{or}\ a
\end{equation}

\subsection{Sensible heat}
This describes the heat loss due to conduction and convection above a hot lake and according to
\citet{Stevenson1992} the ratio between sensible heat, $Q_{sh}$, and heat loss due
to evaporation, $Q_{evap}$ can be written as:

\begin{equation}
    \frac{Q_{sh}}{Q_{evap}}=R_{sh}=\frac{\rho c_a}{L} \frac{(T_s - T_a)}{(e_s - e_a)}
    \underbrace{\left [ \frac{(T_s-T_a)}{(T_{sv} - T_{av})} \right ]^{1/3}}_A 
\end{equation}

where

\begin{tabular}{cp{8cm}}
    $\rho$ & is the air density \\
    $c_a$ & is the specific heat of air\\
\end{tabular}

We decided to ignore term A as it tends to be very close to 1.

\subsection{Putting it all together}
The final equation to compute surface heat losses is then:
\begin{equation}
    Q_e = Q_{rad} + Q_{evap}(1+R_{sh})
\end{equation}

and the mass-loss due to evaporation is:
\begin{equation}
    M_e = Q_{evap}/L
\end{equation}

%\nocite{oreg,schn,pond,smith,marg,hunn,advi,koha,mouse}
